White Dwarf Mergers on Adaptive Meshes\\ II. Collisions

### 2016-07-09 papers/wdmerger_II/paper.tex
Begin the introduction of paper II

>>>>

Binary white dwarf (WD) systems have long been thought to be potential progenitors
of Type Ia supernovae (SNe Ia) \citep{ibentutukov:1984,webbink:1984}. Recently there
has been a growing amount of observational evidence that double WD systems play
an important role in explaining at least some SNe Ia \citep{maoz:2014}. These systems
may be isolated or they may be in hierarchical triple systems, where a WD binary
system in a tight inner orbit is gravitationally coupled to a third, outer star.
In either case the binary WD pair can merge due to the emissions of gravitational waves,
but in the hierarchical triple case, dynamical interactions between the WD binary and
the outer star can prompt mergers as well \citep{thompson:2011,hamers:2013}. The mergers
occur due to perturbations in the eccentricity of the binary orbit caused by the outer
star, which significantly increase the gravitational wave emission of the system. In a
subset of these mergers the eccentricity may be driven to such a high value that the
merger resembles a head-on collision of the binary WDs. Whether the rate of these collisions
may be high enough to significantly contribute to the observed SN Ia rate is uncertain,
though \cite{katzdong:2012} suggest that this is plausible.

Over the past few years a number of groups have performed simulations of WD collisions
to understand whether they may yield astrophysical transients that look like SNe Ia
\citep{rosswog:2009,raskin:2010,loren-aguilar:2010,hawley:2012,garcia-senz:2013,
kushnir:2013,papish:2015,holcomb:2015}. Head-on collisions rapidly convert a significant
amount of kinetic energy into thermal energy in a small region and thus set up the
conditions for a thermonuclear detonation, and these previous simulations have indeed
found that detonations occur and convert a large amount of carbon/oxygen material into
iron-group elements.
